Thesis Proposal

### 2019-09-10 ThesisProposal.Rmd
added proposal requirements and title page

---
title: "Draft Thesis Proposal"
author: "Hanne Oberman"
date: "`r Sys.Date()`"
output: 
  pdf_document: 
    fig_caption: yes
bibliography: ThesisProposal.bib
---

@@

Bayes course on convergence:
```{r}
# E. Assessing convergence
# -----------------------------------------------------------------------------
# History plot, autocorrelation plot

# History plots show the sampled parameters over the iterations (excluding
# the burn-in).
# The development/pattern in these plots gives an indication of convergence.
# When the history plot is stable (a fat catterpillar), convergence is reached.
# Autocorrelation can be measured at many lags.
# High autocorrelation indicates slow mixing of the random path.

mcmcplot(mcmcout = samples)
# #sigma seems fine, but b doesn't:
# #even at lag 5 there is still quite some autocorrelation: therefore we need to center the predictors!
# #otherwise we could use a million iterations to deminish this effect

# Gelman-Rubin diagnostic

# The Gelman and Rubin statistic requires you run
# the sampler/algorithm at least twice: These runs are
# referred to as multiple chains.
# It compares the variance between chains to the variance
# within chains (G-R statistic = T/W = (pooled within chain
# var + between chain var)/pooled within chain var.
# It's the red line in the plot, and should be near 1.
# See Gibbs sampler presentation (week 2)
# https://drive.google.com/file/d/1ABHm8ala3c_puVvf32zF8h6TOZ3898uB/view
# pdf p. 41, slide nr 29.

gelman.plot(samples)
# #this appears to be okay with a really small deviation from 1

# MC error (OPTIONAL?)

# MC error = SD/sqrt(number of iterations)
# SD represents the variation across iterations
# MC error thus represents how much the means differ w.r.t. the iterations
# MC error decreases as number of iterations increases.
# It should not be larger than 5% of the sample standard deviation

@@

# Initial proposal

ShinyMice: an evaluation suite for multiple imputation
Supervisors: Prof.dr. Stef van Buuren, Dr. Gerko Vink

>>>>

---
title: "Thesis Proposal"
subtitle: "Methodology and Statistics for the Behavioural, Biomedical and Social Sciences"
author: "Hanne Oberman"
date: "`r Sys.Date()`"
output: 
  pdf_document: 
    fig_caption: yes
bibliography: ThesisProposal.bib
---

\centering

# ShinyMICE: an Evaluation Suite for Multiple Imputation

Supervised by Prof. Dr. Stef van Buuren, & Dr. Gerko Vink

Department of Methodology and Statistics

Utrecht University

\raggedright

Requirements: max 750 words (excl. references)

In general the proposal should focus on the relevancy of the project, the gap in the scientific literature, and the feasibility to finish the project within 8 months. Possibly specify a 'step 2' to pursue after the main objective is reached.

\newpage

# Proposal by supervisors

@@

"Imputers who do choose to use FCS should use flexible univariate models wherever possible and take care to assess apparent convergence of the algorithm, for example by computing traces of pooled estimates or other statistics and using standard MCMC diagnostics (Gelman et al., 2013, Chpater 11). It may also be helpful to examine the results of many independent runs of the algorithm with different initializations and to use random scans over the p variables to try to identify any convergence issues and mitigate possible order dependence" (@murr18, p. 19).

@@

# Additional resources

Bayes course on convergence:
```{r}
# E. Assessing convergence
# -----------------------------------------------------------------------------
# History plot, autocorrelation plot
# History plots show the sampled parameters over the iterations (excluding
# the burn-in).
# The development/pattern in these plots gives an indication of convergence.
# When the history plot is stable (a fat catterpillar), convergence is reached.
# Autocorrelation can be measured at many lags.
# High autocorrelation indicates slow mixing of the random path.

# mcmcplot(mcmcout = samples)
# #sigma seems fine, but b doesn't:
# #even at lag 5 there is still quite some autocorrelation: therefore we need to center the predictors!
# #otherwise we could use a million iterations to deminish this effect

# Gelman-Rubin diagnostic
# The Gelman and Rubin statistic requires you run
# the sampler/algorithm at least twice: These runs are
# referred to as multiple chains.
# It compares the variance between chains to the variance
# within chains (G-R statistic = T/W = (pooled within chain
# var + between chain var)/pooled within chain var.
# It's the red line in the plot, and should be near 1.
# See Gibbs sampler presentation (week 2)
# https://drive.google.com/file/d/1ABHm8ala3c_puVvf32zF8h6TOZ3898uB/view
# pdf p. 41, slide nr 29.
# gelman.plot(samples)
# #this appears to be okay with a really small deviation from 1

# MC error (OPTIONAL?)
# MC error = SD/sqrt(number of iterations)
# SD represents the variation across iterations
# MC error thus represents how much the means differ w.r.t. the iterations
# MC error decreases as number of iterations increases.
# It should not be larger than 5% of the sample standard deviation
